---
title: "Analysis"
author: "Alexander (Sasha) Pastukhov"
date: "29 4 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggbeeswarm)
library(tidyverse)
```

## Import and preprocessing

Import

```{r import}
data <-
  read_csv("results.csv", 
         col_types = cols(vpcode = col_character(),
                          block = col_integer(),
                          word = col_character(),
                          order = col_character(),
                          response = col_character(),
                          time = col_character())) %>%
  rename(Participant = vpcode) %>%

  
  # extract demographics info
  mutate(Year = as.numeric(stringr::str_sub(Participant, 4, 7)),
         Age = 2021 - Year,
         Sex = stringr::str_sub(Participant, 8, 8),
         Sex = ifelse(Sex == "F", "W", Sex)) # two ways to enter Frau/Weiblich

words <- read_csv("words.csv", 
                  col_types = cols(Word = col_character(),
                                   Condition = col_character(),
                                   Meaning1 = col_character(),
                                   Meaning2 = col_character(),
                                   Meaning2 = col_character()))
```

Converting JSON strings to sequences of values

```{r convert arrays}
# convert JSON entries to atomic vectors
extract_reports <- function(entry){
  tibble(Percept = unlist(str_split(entry$response, ",")) %>% as.numeric(),
         Time = unlist(str_split(entry$time, ",")) %>% as.numeric()) %>%
  mutate(Duration = lead(Time, default=64) - Time)
}

all_reports <-
  data %>%
  group_by(Participant, block, word, Sex, Age) %>%
  nest() %>%
  mutate(Percepts = purrr::map(data, ~extract_reports(.))) %>%
  select(Participant, block, word, Percepts, Sex, Age) %>%
  unnest(Percepts) %>%
  ungroup() %>%
  
  # add word info
  left_join(words, by=c("word" = "Word")) 
```


Extracting individual percepts by concatenating same consecutive responses within the time-series
```{r extract percepts}
all_percepts <-
  all_reports  %>%
  
  # group together repeated reports 
  mutate(iPercept = data.table::rleid(Participant, block, word, Percept)) %>%
  group_by(Participant, block, word, iPercept, Percept, Sex, Age) %>%
  summarise(Time = Time[1], .groups="drop") %>%
  
  # compute duration
  ungroup() %>%
  group_by(Participant, block) %>%
  mutate(Duration = lead(Time) - Time) %>%
  ungroup() %>%

  # compute exact block duration from "end-of-run" percept 50
  group_by(Participant, word) %>%
  mutate(BlockDuration = Time[Percept == 50][1])  %>%

  # add word info
  left_join(words, by=c("word" = "Word")) 
```

Excluding participants with too few reports or who reported no meaning for more than 25% of the time.

```{r valid observers}
participant_summary <-
  all_percepts %>%

  # compute aggregate block indexes
  group_by(Participant, word, Sex, Age) %>%
  summarise(TotalDuration = sum(Duration, na.rm=TRUE),
            NoMeaningDuration = BlockDuration - TotalDuration,
            NoMeaningProportion = NoMeaningDuration / TotalDuration,
            N = n() - 1,
            .groups="drop") %>%
  
  # computing aggregate indexes
  group_by(Participant, Sex, Age) %>%
  summarise(MaxNoMeaningProportion = max(NoMeaningProportion),
            MedianNoMeaningProportion = median(NoMeaningProportion),
            MedianDuration = median(TotalDuration),
            MinResponse = min(N),
            MeanResponses = mean(N),
            MaxResponses = max(N),
            WordsN = length(unique(word)),
            .groups = "drop") %>%
  mutate(IsValid = (WordsN >= nrow(words)-1) & (MedianNoMeaningProportion <= 0.25))

valid_participants <-
  participant_summary %>%
  filter(WordsN == nrow(words),
         MinResponse > 0,
         MedianNoMeaningProportion <= 0.25)

reports <- 
  all_reports %>%
  filter(Participant %in% valid_participants$Participant) 

percepts <-
  all_percepts %>%
  filter(Participant %in% valid_participants$Participant)
```

## Demographics

```{r}
participant_summary %>%
  group_by(Sex) %>%
  summarise(N = n(),
            MinAge = min(Age),
            MaxAge = max(Age),
            ManAge = mean(Age))
```


## Computing aggregate summary per percept

```{r percept summary}
percept_summary <-
  percepts %>%
  group_by(Condition, Participant, Percept, word) %>%
  summarize(N= n(),
            TotalDuration = sum(Duration),
            MeanDuration = mean(Duration),
            Proportion = 100 * sum(Duration) / BlockDuration[1], 
            .groups="drop")
```

## Words with one meaning.
```{r}
unistable <- 
  percept_summary %>%
  filter(Condition == "1") %>%
  
  # filling in zero for percepts that were never reported
  ungroup() %>%
  mutate(PerceptF = factor(Percept, 
                           levels= c(1, -1,  0),
                           labels = c("Main", "Other", "None"))) %>%
  complete(Participant, word, PerceptF, fill=list(Proportion=0, TotalDuration=0, N=0))

unistable_agg <-
  unistable %>%
  filter(PerceptF == "Main") %>%
  group_by(word) %>%
  summarise(Median = median(Proportion),
            LowerQ = quantile(Proportion, (1-0.89)/2),
            UpperQ = quantile(Proportion, 1-(1-0.89)/2),
            .groups="drop") %>%
  mutate(Label = glue::glue("{word}: {round(Median, 1)}%"))

unistable_labels <- unistable_agg$Label
names(unistable_labels) <- unistable_agg$word
  
ggplot(unistable, aes(x=PerceptF, y=Proportion)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_quasirandom(method="tukeyDense", size=0.7) + 
  facet_wrap(.~word, labeller = labeller(word = unistable_labels)) + 
  xlab("Percept") + 
  ylab("Total dominance time [%]")

# ggsave("unistable.png", units="cm", width=14, height=10)
```

## Bistable Words.
```{r}
bistable <- 
  percept_summary %>%
  filter(Condition == "2") %>%

  # filling in zero for percepts that were never reported
  ungroup() %>%
  mutate(PerceptF = factor(Percept, 
                           levels= c("1", "2", "-1",  "0"),
                           labels = c("1", "2", "Other", "None"))) %>%
  complete(Participant, word, PerceptF, fill=list(Proportion=0, MeanDuration=0, N=0))

bistable_agg <-
  bistable %>%
  filter(PerceptF %in% c("1", "2")) %>%
  group_by(Participant, word) %>%
  mutate(Proportion = sum(Proportion)) %>%
  group_by(word) %>%
  summarise(Median = median(Proportion),
            LowerQ = quantile(Proportion, (1-0.89)/2),
            UpperQ = quantile(Proportion, 1-(1-0.89)/2),
            .groups="drop") %>%
  mutate(Label = glue::glue("{word}: {round(Median, 1)}%"))

bistable_labels <- bistable_agg$Label
names(bistable_labels) <- bistable_agg$word

  
ggplot(bistable, aes(x=PerceptF, y=Proportion)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_quasirandom(method="tukeyDense", size=0.7) + 
  facet_wrap(.~word, labeller = labeller(word = bistable_labels), ncol = 3) + 
  xlab("Percept") + 
  ylab("Total dominance time [%]")
```
Computing balance between main meanings

```{r balance}
meaning_balance <-
  bistable %>%
  filter(Percept %in% c(1, 2)) %>%
  group_by(Participant, word) %>%
  summarize(JointProportion = sum(Proportion),
            MeanProportion = median(Proportion),
            Balance = Proportion[1] / sum(Proportion),
            dProportion = abs(diff(Proportion)), 
            .groups="drop") 

meaning_balance_agg <-
  meaning_balance %>%
  group_by(word) %>%
  summarise(MeanProportionSD = round(sd(MeanProportion), 2),
            MeanProportionLo = quantile(MeanProportion, (1-0.89)/2),
            MeanProportionHi = quantile(MeanProportion, 1-(1-0.89)/2),
            MeanProportion = round(mean(MeanProportion), 2),
            JointProportionSD = round(sd(JointProportion), 2),
            JointProportion = round(median(JointProportion), 2),
            
            MeanBalance = mean(Balance),
            BalanceSE = sd(Balance) / sqrt(n()-1),
            BalanceLo = quantile(Balance, (1-0.89)/2),
            BalanceHi = quantile(Balance, 1-(1-0.89)/2),
            
            dProportionSD = round(sd(dProportion), 2),
            dProportion = round(median(dProportion), 2)) %>%
  ungroup()  %>%
  relocate(word, MeanProportion, MeanProportionSD, dProportion, dProportionSD, JointProportion, JointProportionSD) %>%
  arrange(dProportion, desc(MeanProportion), desc(JointProportion))

best_balance <- meaning_balance_agg$word[1:4]

meaning_balance_agg
```

```{r}
ggplot(meaning_balance_agg, aes(x=word, y=MeanBalance, ymin=MeanBalance-BalanceSE, ymax=MeanBalance+BalanceSE)) +  # ymin=BalanceLo, ymax=BalanceHi)) + 
  geom_errorbar() + 
  geom_hline(yintercept = 0.5, color="white", size=1) +
  geom_point() + 
  xlab(NULL) + 
  ylab("Balance [mean ± SE]") + 
  # scale_y_continuous(limits=c(0, 50)) 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggsave("bistable-balance.png", units="cm", width=16, height=8)

```


Time series for most balanced words
```{r}
best_balance <- c("Absatz", "Kiefer", "Schloss", "Gericht")

ggplot(reports %>% 
         filter(Percept != 50, word %in% best_balance[1], !Participant %in% c("CSO2000W", "GML1998W", "XNB2001W", "GEN1985M", "NBF1996W") ) %>%
         mutate(Percept = ifelse(Percept == -1, 1.5, Percept)),
       aes(x=Time, y=Percept)) + 
  geom_line() + #aes(color=Participant), show.legend = FALSE) +
  facet_wrap(.~Participant) + 
  scale_y_continuous(breaks = c(1, 1.5, 2), labels = c("1", "other", "2")) + 
  scale_x_continuous("Time [s]", limits = c(0, 65))

ggsave("absatz-timeseries.png", units="cm", width=16, height=10)
```

### Tristable words

```{r}
tristable <- 
  percept_summary %>%
  filter(Condition == "3") %>%

  # filling in zero for percepts that were never reported
  ungroup() %>%
  mutate(PerceptF = factor(Percept, 
                           levels= c(1, 2, 3, -1,  0),
                           labels = c("1", "2", "3", "Other", "None"))) %>%
  complete(Participant, word, PerceptF, fill=list(Proportion=0))

tristable_agg <-
  tristable %>%
  filter(PerceptF %in% c("1", "2", "3")) %>%
  group_by(Participant, word) %>%
  mutate(Proportion = sum(Proportion)) %>%
  group_by(word) %>%
  summarise(Median = median(Proportion),
            LowerQ = quantile(Proportion, (1-0.89)/2),
            UpperQ = quantile(Proportion, 1-(1-0.89)/2),
            .groups="drop") %>%
  mutate(Label = glue::glue("{word}: {round(Median, 1)}%"))

tristable_labels <- tristable_agg$Label
names(tristable_labels) <- tristable_agg$word


ggplot(tristable, aes(x=PerceptF, y=Proportion)) + 
geom_boxplot(outlier.shape = NA) +
  geom_quasirandom(method="tukeyDense", size=0.7) + 
  facet_wrap(.~word, labeller = labeller(word = tristable_labels), ncol = 3) + 
  xlab("Percept") + 
  ylab("Total dominance time [%]")


ggsave("tristable.png", units="cm", width=14, height=10)
```


