---
title: "Analysis"
author: "Alexander (Sasha) Pastukhov"
date: "29 4 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)

library(binom)
library(ggbeeswarm)
library(patchwork)
library(tidyverse)
```

## Utilities

```{r}
custom_plot_theme <- function()
{
  theme(# text=element_text(family="Arial"),
        axis.text.x = element_text(size = 8, colour = 'black'), 
        axis.text.y = element_text(size = 8, colour = 'black'), 
        axis.title.x = element_text(size = 10), 
        axis.title.y = element_text(size = 10), 
        panel.grid.minor.x = element_blank(), 
        panel.grid.minor.y =  element_line(size = 0.24), 
        axis.ticks = element_line(size = 0.24),
        plot.subtitle=element_text(size=8, hjust=0.5),
        strip.text = element_text(size = 8))
}
```

## Import and preprocessing

Import

```{r import}
data <-
  read_csv("results.csv", 
         col_types = cols(vpcode = col_character(),
                          block = col_integer(),
                          word = col_character(),
                          order = col_character(),
                          response = col_character(),
                          time = col_character())) %>%
  rename(Participant = vpcode) %>%

  
  # extract demographics info
  mutate(Year = as.numeric(stringr::str_sub(Participant, 4, 7)),
         Age = 2021 - Year,
         Sex = stringr::str_sub(Participant, 8, 8),
         Sex = ifelse(Sex == "F", "W", Sex)) # two ways to enter Frau/Weiblich

words <- read_csv("words.csv", 
                  col_types = cols(Word = col_character(),
                                   Condition = col_character(),
                                   Meaning1 = col_character(),
                                   Meaning2 = col_character(),
                                   Meaning2 = col_character()))
```

Converting JSON strings to sequences of values

```{r convert arrays}
# convert JSON entries to atomic vectors
extract_reports <- function(entry){
  tibble(Percept = unlist(str_split(entry$response, ",")) %>% as.numeric(),
         Time = unlist(str_split(entry$time, ",")) %>% as.numeric()) %>%
  mutate(Duration = lead(Time, default=64) - Time)
}

all_reports <-
  data %>%
  group_by(Participant, block, word, Sex, Age) %>%
  nest() %>%
  mutate(Percepts = purrr::map(data, ~extract_reports(.))) %>%
  select(Participant, block, word, Percepts, Sex, Age) %>%
  unnest(Percepts) %>%
  ungroup() %>%
  
  # add word info
  left_join(words, by=c("word" = "Word")) 
```


Extracting individual percepts by concatenating same consecutive responses within the time-series
```{r extract percepts}
all_percepts <-
  all_reports  %>%
  
  # group together repeated reports 
  mutate(iPercept = data.table::rleid(Participant, block, word, Percept)) %>%
  group_by(Participant, block, word, iPercept, Percept, Sex, Age) %>%
  summarise(Time = Time[1], .groups="drop") %>%
  
  # compute duration
  ungroup() %>%
  group_by(Participant, block) %>%
  mutate(Duration = lead(Time) - Time) %>%
  ungroup() %>%

  # compute exact block duration from "end-of-run" percept 50
  group_by(Participant, word) %>%
  mutate(BlockDuration = Time[Percept == 50][1])  %>%

  # add word info
  left_join(words, by=c("word" = "Word")) %>%
  
  # drop end-of-block Percept == 50
  filter(Percept != 50)
```

Excluding participants with too few reports or who reported no meaning for more than 25% of the time.

```{r valid observers}
participant_summary <-
  all_percepts %>%

  # compute aggregate block indexes
  group_by(Participant, word, Sex, Age) %>%
  summarise(TotalDuration = sum(Duration, na.rm=TRUE),
            NoMeaningDuration = BlockDuration - TotalDuration,
            NoMeaningProportion = NoMeaningDuration / TotalDuration,
            N = n() - 1,
            .groups="drop") %>%
  
  # computing aggregate indexes
  group_by(Participant, Sex, Age) %>%
  summarise(MaxNoMeaningProportion = max(NoMeaningProportion),
            MedianNoMeaningProportion = median(NoMeaningProportion),
            MedianDuration = median(TotalDuration),
            MinResponse = min(N),
            MeanResponses = mean(N),
            MaxResponses = max(N),
            WordsN = length(unique(word)),
            .groups = "drop") %>%
  mutate(IsValid = (WordsN >= nrow(words)-1) & (MedianNoMeaningProportion <= 0.25))

valid_participants <-
  participant_summary %>%
  filter(WordsN == nrow(words),
         MedianNoMeaningProportion <= 0.25)

reports <- 
  all_reports %>%
  filter(Participant %in% valid_participants$Participant) 

percepts <-
  all_percepts %>%
  filter(Participant %in% valid_participants$Participant)
```

## Demographics
All participants
```{r}
participant_summary %>%
  group_by(Sex) %>%
  summarise(N = n(),
            MinAge = min(Age),
            MaxAge = max(Age),
            MeanAge = mean(Age))
```

Valid participants (final sample)
```{r}
valid_participants %>%
  group_by(Sex) %>%
  summarise(N = n(),
            MinAge = min(Age),
            MaxAge = max(Age),
            MeanAge = mean(Age))
```


## Computing aggregate summary per percept

```{r percept summary}
percept_summary <-
  percepts %>%
  group_by(Condition, Participant, Percept, word) %>%
  summarize(N= n(),
            TotalDuration = sum(Duration, na.rm=TRUE),
            MeanDuration = mean(Duration, na.rm=TRUE),
            Proportion = 100 * sum(Duration) / BlockDuration[1], 
            .groups="drop")
```

## Words with one meaning.
```{r}
unistable <- 
  percept_summary %>%
  filter(Condition == "1") %>%
  
  # filling in zero for percepts that were never reported
  ungroup() %>%
  mutate(PerceptF = factor(Percept, 
                           levels= c(1, -1,  0),
                           labels = c("Main", "Other", "None"))) %>%
  complete(Participant, word, PerceptF, fill=list(Proportion=0, TotalDuration=0, N=0))

unistable_agg <-
  unistable %>%
  filter(PerceptF == "Main") %>%
  group_by(word) %>%
  summarise(Median = median(Proportion),
            LowerQ = quantile(Proportion, 0.25),
            UpperQ = quantile(Proportion, 0.75),
            .groups="drop") %>%
  mutate(Label = glue::glue("{word}\n{round(Median, 1)}%, [{round(LowerQ, 1)}%..{round(UpperQ, 1)}%]"))

unistable_labels <- unistable_agg$Label
names(unistable_labels) <- unistable_agg$word
  
ggplot(unistable, aes(x=PerceptF, y=Proportion)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_quasirandom(method="tukeyDense", size=0.7, width=0.1) + 
  facet_wrap(.~word, labeller = labeller(word = unistable_labels)) + 
  xlab("Percept") + 
  ylab("Total dominance time [%]") + 
  custom_plot_theme()
  

# ggsave("unistable.png", units="cm", width=14, height=10)
# ggsave("unistable.svg", units="cm", width=14, height=10)
```


## Bistable Words.
```{r}
bistable <- 
  percept_summary %>%
  filter(Condition == "2") %>%

  # filling in zero for percepts that were never reported
  ungroup() %>%
  mutate(PerceptF = factor(Percept, 
                           levels= c("1", "2", "-1",  "0"),
                           labels = c("1", "2", "Other", "None"))) %>%
  complete(Participant, word, PerceptF, fill=list(Proportion=0, MeanDuration=0, N=0))

bistable_agg <-
  bistable %>%
  filter(PerceptF %in% c("1", "2")) %>%
  group_by(Participant, word) %>%
  mutate(Proportion = sum(Proportion)) %>%
  group_by(word) %>%
  summarise(Median = median(Proportion),
            LowerQ = quantile(Proportion, 0.25),
            UpperQ = quantile(Proportion, 0.75),
            .groups="drop") %>%
  mutate(Label = glue::glue("{word}\n {round(Median, 1)}%, [{round(LowerQ, 1)}%..{round(UpperQ, 1)}%]"))

bistable_labels <- bistable_agg$Label
names(bistable_labels) <- bistable_agg$word

  
ggplot(bistable, aes(x=PerceptF, y=Proportion)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_quasirandom(method="tukeyDense", size=0.7, width=0.1) + 
  facet_wrap(.~word, labeller = labeller(word = bistable_labels), ncol = 3) + 
  custom_plot_theme() +
  xlab("Percept") + 
  ylab("Total dominance time [%]")

# ggsave("bistable.png", units="cm", width=14, height=20)
# ggsave("bistable.svg", units="cm", width=14, height=20)
```

Computing balance between main meanings and  probability that meaning #1 came first
```{r}
bistable_balance_and_order <-
  percepts %>%
  
  # bistable words, primary meanings only
  filter(Condition == 2, Percept %in% 1:2) %>%
  
  group_by(Participant, word) %>%
  summarise(FirstMeaning = Percept[1], 
            TotalDuration = sum(Duration),
            Duration1 = sum(Duration[Percept==1]),
            Duration2 = sum(Duration[Percept==2]),
            Balance = Duration1 / TotalDuration,
            .groups="drop")

balance_plot <-
  ggplot(bistable_balance_and_order, aes(x=word, y=Balance)) + 
  geom_boxplot(outlier.shape = NA) + 
  geom_hline(yintercept = 0.5, color="white", size=1) +
  geom_quasirandom(method="tukeyDense", width=0.1, size=0.7) + 
  xlab(NULL) + 
  ylab("Balance of meanings") + 
  custom_plot_theme() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))



bistable_order <- 
  bistable_balance_and_order %>%
  group_by(word) %>%
  summarise(Bmedian = median(Balance),
            Blower = quantile(Balance, 0.25),
            Bupper = quantile(Balance, 0.75),
            Pfirst = mean(FirstMeaning == 1),
            Lfirst = binom::binom.exact(sum(FirstMeaning == 1), n())$lower,
            Ufirst = binom::binom.exact(sum(FirstMeaning == 1), n())$upper,
            .groups="drop")

order_plot <- 
  ggplot(bistable_order, aes(x=word, y=Pfirst, ymin=Lfirst, ymax=Ufirst)) + 
  geom_point() +
  geom_errorbar() +
  geom_hline(yintercept = 0.5, color="white", size=1) +
  xlab(NULL) +
  ylab("P(Report[1]=meaning #1)") + 
  custom_plot_theme() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

balance_order_cor <- cor.test(bistable_order$Bmedian, bistable_order$Pfirst)
balance_order_cor_text <- glue::glue("ρ={round(balance_order_cor$estimate, 2)}, 95% CI = {round(balance_order_cor$conf.int[1], 2)}..{round(balance_order_cor$conf.int[2], 2)}")

corr_plot <- 
  ggplot(bistable_order, aes(x=Bmedian, y=Pfirst)) + 
  geom_smooth(method="lm", formula=y~x) +
  geom_label(aes(label=word)) +
  xlim(0.25, 1.1) + 
  ylim(0.25, 1.1) + 
  xlab("Balance of meanings") + 
  ylab("P(Report[1]=meaning #1)") +
  labs(subtitle=balance_order_cor_text) +
  coord_equal() +
  custom_plot_theme()

all_plots <- (balance_plot | order_plot) / corr_plot
all_plots + plot_annotation(tag_levels = 'A') + plot_layout(heights=c(1, 2.5))

# ggsave("bistable-balance-order.png", units="cm", width=14, height=16)
# ggsave("bistable-balance-order.svg", units="cm", width=14, height=16)
```



Time series for most balanced words
```{r}
best_balance <- c("Absatz", "Kiefer", "Schloss", "Gericht")


x <- reports %>% 
         filter(Percept != 50, word %in% best_balance[1]) %>%
         mutate(Percept = ifelse(Percept == 2, 3, Percept),
                Percept = ifelse(Percept == -1, 2, Percept))
                
ggplot(reports %>% 
         filter(Percept != 50, word %in% best_balance[1]) %>%
         mutate(Percept = ifelse(Percept == 2, 3, Percept)) %>%
         mutate(Percept = ifelse(Percept == -1, 2, Percept)),
       aes(x=Time, y=Percept)) + 
  geom_line() + #aes(color=Participant), show.legend = FALSE) +
  facet_wrap(.~Participant) + 
  scale_y_continuous(NULL, breaks = c(0, 1, 2, 3), labels = c("None", "Stöckel", "Other meaning", "Abschnitt")) + 
  scale_x_continuous("Time [s]", limits = c(0, 65)) + 
  custom_plot_theme()

# ggsave("absatz-timeseries.png", units="cm", width=16, height=14)
# ggsave("absatz-timeseries.svg", units="cm", width=16, height=14)
```

### Tristable words

```{r}
tristable <- 
  percept_summary %>%
  filter(Condition == "3") %>%

  # filling in zero for percepts that were never reported
  ungroup() %>%
  mutate(PerceptF = factor(Percept, 
                           levels= c(1, 2, 3, -1,  0),
                           labels = c("1", "2", "3", "Other", "None"))) %>%
  complete(Participant, word, PerceptF, fill=list(Proportion=0))

tristable_agg <-
  tristable %>%
  filter(PerceptF %in% c("1", "2", "3")) %>%
  group_by(Participant, word) %>%
  mutate(Proportion = sum(Proportion)) %>%
  group_by(word) %>%
  summarise(Median = median(Proportion),
            LowerQ = quantile(Proportion, 0.25),
            UpperQ = quantile(Proportion, 0.75),
            .groups="drop") %>%
  mutate(Label = glue::glue("{word}\n {round(Median, 1)}%, [{round(LowerQ, 1)}%..{round(UpperQ, 1)}%]"))

tristable_labels <- tristable_agg$Label
names(tristable_labels) <- tristable_agg$word


ggplot(tristable, aes(x=PerceptF, y=Proportion)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_quasirandom(method="tukeyDense", size=0.7) + 
  facet_wrap(.~word, labeller = labeller(word = tristable_labels), ncol = 3) + 
  xlab("Percept") + 
  ylab("Total dominance time [%]")


# ggsave("tristable.png", units="cm", width=14, height=10)
# ggsave("tristable.png", units="cm", width=14, height=10)
```

## Combined figure for iPerception paper


Unistable
```{r}
uniplot <-
ggplot(filter(unistable, word %in% c("Brief", "Lampe")), aes(x=PerceptF, y=Proportion)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_quasirandom(method="tukeyDense", size=0.7, width=0.1) + 
  facet_wrap(.~word, labeller = labeller(word = unistable_labels)) + 
  xlab("Percept") + 
  scale_y_continuous(name = "Total dominance time [%]", limits = c(0, 100)) +
  custom_plot_theme()
```


Bistable
```{r}
biplot <-
ggplot(filter(bistable, word %in% c("Absatz", "Ball")), aes(x=PerceptF, y=Proportion)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_quasirandom(method="tukeyDense", size=0.7, width=0.1) + 
  facet_wrap(.~word, labeller = labeller(word = bistable_labels), ncol = 3) + 
  custom_plot_theme() +
  xlab("Percept") + 
  scale_y_continuous(name = "Total dominance time [%]", limits = c(0, 100))

```

Tristable
```{r}
triplot <-
ggplot(filter(tristable, word %in% c("Decken", "Mine")), aes(x=PerceptF, y=Proportion)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_quasirandom(method="tukeyDense", size=0.7) + 
  facet_wrap(.~word, labeller = labeller(word = tristable_labels), ncol = 3) + 
  xlab("Percept") + 
  scale_y_continuous(name = "Total dominance time [%]", limits = c(0, 100)) +
  custom_plot_theme()
```

Bistable words balance
```{r}
order_plot <-
  ggplot(bistable_order, aes(x=word, y=Pfirst, ymin=Lfirst, ymax=Ufirst)) + 
  geom_point() +
  geom_errorbar() +
  geom_hline(yintercept = 0.5, color="white", size=1) +
  xlab(NULL) +
  ylab("P(Report[1]=meaning #1)") + 
  custom_plot_theme() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


balance_plot <-
bistable_balance_and_order |>
  mutate(word = fct_reorder(word, Balance)) |>
  ggplot(aes(x=word, y=Balance)) + 
  geom_boxplot(outlier.shape = NA) + 
  geom_hline(yintercept = 0.5, color="white", size=1) +
  geom_quasirandom(method="tukeyDense", width=0.1, size=0.7) + 
  xlab(NULL) + 
  ylab("Balance of meanings") + 
  custom_plot_theme() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```


```{r}
timeseriesplot <-
  ggplot(reports %>% 
         filter(Percept != 50, word %in% best_balance[1]) %>%
         filter(Participant %in% c("BAW1995M", "BKB2002W", "BTS2001W", "KDD1998W", "CSO2000W", "UDA1993F")) |>
         mutate(Percept = ifelse(Percept == 2, 3, Percept)) %>%
         mutate(Percept = ifelse(Percept == -1, 2, Percept)),
       aes(x=Time, y=Percept)) + 
  geom_line() + #aes(color=Participant), show.legend = FALSE) +
  facet_wrap(.~Participant) + 
  scale_y_continuous(NULL, breaks = c(0, 1, 2, 3), labels = c("None", "Stöckel", "Other", "Abschnitt")) +
  scale_x_continuous("Time [s]", limits = c(0, 65)) + 
  custom_plot_theme()
```

```{r}
(uniplot / biplot / triplot) | (balance_plot / timeseriesplot) +
  plot_layout(widths = c(1.5, 1))
```
